Multilocus Genotype And Mycelial Compatability Group Assessment

### 2017-08-14 doc/RMD/mlg-mcg.Rmd
Update figures to fix #9
Ironically, this doesn't explicitly simplify, it's now a bit more complex in content! :joy_cat:

I've also started saving descriptive figure names.
At some point I will burn all the figures down and only the descriptive names will be spared.

```{r barplots, fig.width = 10, fig.height = 5}
mcg_graph <- mcgs %>% 
  filter(Samples > 1) %>%
  gather(type, count, MLGs, Samples, -Evenness) %>%
  arrange(desc(type), desc(count)) %>%
  rename(Type = type) %>%
  mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
  ggplot(aes(x = MCG, y = count, group = Type, fill = Type, alpha = Evenness)) +
  geom_col(aes(width = ifelse(Type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
  # annotate(geom = "text", x = 18, y =35, 
  #          label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
  scale_fill_manual(values = c("black", "white")) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal(base_size = 16, base_family = "Helvetica") +
  # coord_flip() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.major.y = element_line(color = "grey50")) +
  theme(panel.grid.minor.y = element_line(color = "grey70")) +
  theme(legend.box.just = "top") +
  theme(legend.box.margin = unit(c(0, 6.55, 0, 0), "lines")) +
  theme(legend.box.background = element_rect(fill = "white", color = "black")) +
  theme(legend.box = "horizontal") +
  theme(aspect.ratio = 1/2) +
  labs(list(
    alpha = "MLG\nEvenness"
  ))

@@

vp1 <- grid::viewport(width = 0.25, height = 0.4, x = 1, y = 0.975, just = c("right", "top"))
mcg_graph <- mcg_graph + theme(legend.position = c(.763, .75))

print(mcg_graph)
print(even, vp = vp1)

if (!interactive()){
  pdf(file.path(PROJHOME, "results/figures/publication/Figure1Z.pdf"), width = 10, height = 5)
  print(mcg_graph)
  print(even, vp = vp1)
  dev.off()
}
```

@@

make_mcgmlg_graph <- function(x){
  gdf <- mutate(x, MLG = paste0('MLG.', MLG))
  MLGS <- gdf %>% 
    group_by(MLG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MLG)
  MCGS <- gdf %>% 
    group_by(MCG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MCG)
  VAT <- bind_rows(MLGS, MCGS)
  g <- gdf %>% 
    select(MCG, MLG, Freq) %>%
    rename(weight = Freq) %>%
    graph_from_data_frame(vertices = VAT)
  V(g)$type <- ifelse(grepl("MLG", V(g)$name), "Multilocus Genotype", "Mycelial Compatibility Group")
  g
}

@@

the_communities <- data_frame(vertex = lay2$name, community = lay2$community) %>%
  mutate(comm = case_when(
    .$community == 7 ~ "A",
    .$community == 5 ~ "B",
    .$community == 1 ~ "C",
    TRUE ~ as.character(.$community)
  )) %>%
  group_by(community) %>%
  mutate(comm = ifelse(n() > 10, paste("Community", comm), "Other Communities (n < 10)")) %>%
  mutate(comm = comm) # It just so happens that the communities are in order by size.
lay2$community <- the_communities$comm
mcg_mlg_graph <- ggraph(lay2) +
  geom_node_circle(aes(r = size, lty = type, fill = community)) +
  geom_edge_link(aes(start_cap = circle(node1.size, unit = "native"), 
                     end_cap = circle(node2.size, unit = "native"), 
                     width = weight),
                 arrow = arrow(length = unit(0.01, "native"))) +
  geom_node_text(aes(label = gsub("MLG.", "", name), 
                     color = community, 
                     size = size/10), 
                 show.legend = FALSE) +
  coord_fixed() +
  scale_linetype_manual(values = c(1, 3)) +
  scale_edge_width(range = c(0.25, 1.5)) +
  scale_color_manual(values = c("black", "black", "white", "white")) +
  scale_fill_viridis(direction = -1, discrete = TRUE) +
  theme_graph(base_size = 16, base_family = "Helvetica") +
  theme(legend.position = "bottom", legend.direction = "vertical") +
  theme(legend.box.spacing = unit(-1, "lines")) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "lines"))
  # ggtitle("Relation of Multilocus Genotypes and MCGs")

mcg_mlg_graph
if (!interactive()){
  ggsave(plot = mcg_mlg_graph, filename = file.path(PROJHOME, "results/figures/publication/FigureS2.pdf"), 
       width = 88*3, height = 88*3.5, units = "mm")
  tmp <- new.env()
  assign("mcg_mlg_igraph", g, envir = tmp)
  assign("mcg_mlg_gg", mcg_mlg_graph, envir = tmp)
  V(tmp$mcg_mlg_igraph)$size <- osize
  save(list = ls(envir = tmp), 
       envir = tmp, 
       file = file.path(PROJHOME, "data", "mcg-mlg-graph.rda"))
}

cat("Community Structure:\n")
cluster_walktrap(g)
```

@@

So that's a big hairy graph. What happens when we look at a subgraph of the
top 5 MCGs?

```{r top5graph, fig.height = 7, fig.width = 7}
top5 <- filter(mcgmlg, as.character(MCG) %in% mcgs$MCG[1:5])
top5g <- make_mcgmlg_graph(top5)
tosize <- V(top5g)$size
V(top5g)$size <- sqrt(tosize)/10
set.seed(2017-05-08)
top5lay <- create_layout(top5g, layout = "igraph", algorithm = "fr", weights = rep(1, length(E(top5g))))
top5lay$community <- inner_join(data_frame(vertex = top5lay$name), the_communities)$comm
t5g <- mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
t5g
if (!interactive())
  ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.pdf"), width = 88, height = 88, scale = 3, units = "mm")
```

@@

>>>>

```{r barplots, fig.width = 10, fig.height = 5}
mcg_graph <- mcgs %>% 
  filter(Samples > 1) %>%
  gather(type, count, MLGs, Samples, -Evenness) %>%
  arrange(desc(type), desc(count)) %>%
  rename(Type = type) %>%
  mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
  ggplot(aes(x = MCG, y = count, group = Type, fill = Type, alpha = Evenness)) +
  geom_col(aes(width = ifelse(Type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
  # annotate(geom = "text", x = 18, y =35, 
  #          label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
  scale_fill_manual(values = c("black", "white")) +
  scale_y_continuous(expand = c(0, 2)) +
  # scale_alpha(breaks = c(0.5, 0.75, 1)) +
  theme_minimal(base_size = 16, base_family = "Helvetica") +
  # coord_flip() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.major.y = element_line(color = "grey50")) +
  theme(panel.grid.minor.y = element_line(color = "grey70")) +
  theme(legend.box.just = "top") +
  theme(legend.box.margin = unit(c(0, 6.55, 0, 0), "lines")) +
  theme(legend.box.background = element_rect(fill = "white", color = "black")) +
  theme(legend.box = "horizontal") +
  theme(aspect.ratio = 1/2) +
  labs(list(
    alpha = "MLG\nEvenness"
  ))

@@

# vp1 <- grid::viewport(width = 0.25, height = 0.4, x = 1, y = 0.975, just = c("right", "top"))
# mcg_graph <- mcg_graph + theme(legend.position = c(.763, .75))
# print(mcg_graph)
# print(even, vp = vp1)

mcg_graph_inset <- cowplot::ggdraw() + 
  cowplot::draw_plot(mcg_graph, 0, 0, 1, 1) + 
  cowplot::draw_plot(even, 0.75, 0.5675, 0.25, 0.4)
mcg_graph_inset
if (!interactive()){
  cowplot::ggsave(mcg_graph_inset, 
                  filename = file.path(PROJHOME, "results/figures/publication/mcg-barplot.pdf"),
                  width = 10,
                  height = 5)
  # pdf(file.path(PROJHOME, "results/figures/publication/Figure1Z.pdf"), width = 10, height = 5)
  # print(mcg_graph)
  # print(even, vp = vp1)
  # dev.off()
}
```

@@

make_mcgmlg_graph <- function(x){
  gdf <- mutate(x, MLG = paste0('MLG.', MLG))
  MLGS <- gdf %>% 
    group_by(MLG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MLG)
  MCGS <- gdf %>% 
    group_by(MCG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MCG)
  VAT <- bind_rows(MLGS, MCGS)
  g <- gdf %>% 
    select(MLG, MCG, Freq) %>%
    rename(weight = Freq) %>%
    graph_from_data_frame(vertices = VAT)
  V(g)$type <- ifelse(grepl("MLG", V(g)$name), "Multilocus Genotype", "Mycelial Compatibility Group")
  g
}

@@

the_communities <- data_frame(vertex = lay2$name, community = lay2$community) %>%
  mutate(comm = case_when(
    .$community == 7 ~ "A",
    .$community == 5 ~ "B",
    .$community == 1 ~ "C",
    TRUE ~ as.character(.$community)
  )) %>%
  group_by(community) %>%
  mutate(comm = ifelse(n() > 10, paste("Community", comm), "Other Communities (n < 10)")) %>%
  mutate(comm = comm) # It just so happens that the communities are in order by size.
lay2$community <- the_communities$comm
mcg_mlg_graph <- ggraph(lay2) +
  geom_node_circle(aes(r = size, lty = type, fill = community)) +
  geom_edge_link(aes(start_cap = circle(node1.size, unit = "native"), 
                     end_cap = circle(node2.size, unit = "native"), 
                     width = weight),
                 arrow = arrow(length = unit(0.005, "native"))) +
  geom_node_text(aes(label = gsub("MLG.", "", name), 
                     color = community, 
                     size = size/10), 
                 show.legend = FALSE) +
  coord_fixed() +
  scale_linetype_manual(values = c(1, 3)) +
  scale_edge_width(range = c(0.25, 1.5)) +
  scale_color_manual(values = c("black", "black", "white", "white")) +
  scale_fill_viridis(direction = -1, discrete = TRUE) +
  theme_graph(base_size = 16, base_family = "Helvetica") +
  theme(legend.position = "bottom", legend.direction = "vertical") +
  theme(legend.box.spacing = unit(-1, "lines")) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "lines"))
  # ggtitle("Relation of Multilocus Genotypes and MCGs")

mcg_mlg_graph
if (!interactive()){
  ggsave(plot = mcg_mlg_graph, filename = file.path(PROJHOME, "results/figures/publication/full-graph.pdf"), 
       width = 88*3, height = 88*3.5, units = "mm")
  tmp <- new.env()
  assign("mcg_mlg_igraph", g, envir = tmp)
  assign("mcg_mlg_gg", mcg_mlg_graph, envir = tmp)
  V(tmp$mcg_mlg_igraph)$size <- osize
  save(list = ls(envir = tmp), 
       envir = tmp, 
       file = file.path(PROJHOME, "data", "mcg-mlg-graph.rda"))
}
cat("Community Structure:\n")
cluster_walktrap(g)
```

Because showing this big, hairy graph can overload a reader, I'm creating an
inset graph for later use.

```{r mlgmcggraph2, fig.width = 2, fig.height = 2}
mcg_mlg_graph2 <- ggraph(lay2) +
  geom_edge_arc(#aes(width = weight), 
                start_cap = circle(1.1, "mm"), 
                end_cap = circle(1.1, "mm"), 
                curvature = 0.1
                ) + #arrow = arrow(length = unit(2, "mm"))) +
  geom_node_point(aes(fill = type), pch = 21, size = 0.75) +
  coord_fixed() +
  scale_edge_width(range = c(0.25, 1.5)) +
  scale_fill_manual(values = c("black", "white")) +
  theme_graph(base_size = 16, base_family = "Helvetica") +
  theme(legend.position = "none", legend.direction = "vertical") +
  # theme(legend.box.spacing = unit(-1, "lines")) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "lines")) +
  theme(plot.background = element_rect(fill = "grey98", color = "black"))
mcg_mlg_graph2
```

@@

So that's a big hairy graph. What happens when we look at a subgraph of the
top 8 MCGs? Note, that these are all MCGs that have > 10 isolates associated.

```{r top5graph, fig.height = 7, fig.width = 7}
top5 <- filter(mcgmlg, as.character(MCG) %in% (mcgs %>% arrange(desc(Samples)) %>% slice(1:8) %>% pull(MCG)))
top5g <- make_mcgmlg_graph(top5)
tosize <- V(top5g)$size
V(top5g)$size <- sqrt(tosize)/10
# set.seed(2017-05-08)
set.seed(2017-08-1)
top5lay <- create_layout(top5g, layout = "igraph", algorithm = "fr", weights = rep(1, length(E(top5g))))
top5lay$community <- inner_join(data_frame(vertex = top5lay$name), the_communities)$comm
levels(top5lay$type) <- c("Multilocus Genotype", "Mycelial Compatibility\nGroup")
t5g <- ggraph(top5lay) +
  geom_node_circle(aes(r = size, fill = type)) +
  geom_edge_link(aes(start_cap = circle(node1.size, unit = "native"), 
                     end_cap = circle(node2.size, unit = "native"), 
                     width = weight),
                 arrow = arrow(length = unit(0.01, "native"))) +
  geom_node_text(aes(label = gsub("MLG.", "", name), 
                     color = type, 
                     size = size/10),
                 show.legend = FALSE) +
  coord_fixed() +
  scale_edge_width(range = c(0.25, 1.5)) +
  scale_fill_manual(values = c("black", "white")) +
  scale_color_manual(values = c("white", "black")) +
  theme_graph(base_size = 16, base_family = "Helvetica") +
  # theme(legend.position = c(0.85, 0.35), legend.direction = "vertical") +
  theme(legend.position = c(0.5, 0), legend.direction = "horizontal", legend.box = "horizontal") +
  theme(legend.box.spacing = unit(-1, "lines")) +
  theme(legend.spacing.x = unit(0, "lines")) +
  # theme(legend.box.background = element_rect(color = "black")) +
  theme(legend.text = element_text(size = 10)) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "lines"))
t5g$labels$edge_width <- "shared isolates"
t5g$labels$fill <- ""
t5g

if (!interactive())
  ggsave(t5g, file = file.path("results", "figures", "publication", "mcg-mlg-subgraph.pdf"), width = 88, height = 88, scale = 3, units = "mm")
```

@@

```{r subgraphs, fig.height = 10, fig.width = 8}
mcg_theme <- theme(legend.text = element_text(size = 10)) +
  theme(legend.title = element_text(size = 10)) +
  theme(axis.text.x = element_blank()) +
  # theme(legend.background = element_blank()) +
  # theme(legend.box.background = element_blank()) +
  theme(legend.key.size = unit(0.5, "line")) +
  theme(legend.position = c(0.9, 0.5)) +
  theme(legend.box = "vertical") +
  theme(legend.box.margin = unit(c(0, 0, 0, 0), "lines")) +
  theme(legend.spacing = unit(0, "lines")) + 
  theme(aspect.ratio = 0.45) +
  theme(plot.background = element_rect(color = "black", fill = "grey98"))
mcg_subgraph_with_context <- cowplot::ggdraw(xlim = c(0, 2), ylim = c(0, 2.25)) +
  cowplot::draw_plot(t5g, 0, -0.125, 2, 2) +
  cowplot::draw_plot(mcg_graph + scale_alpha(breaks = c(0.5, 0.75, 1)) + mcg_theme, 0, 1.45, 1.35, 1) +
  cowplot::draw_plot(mcg_mlg_graph2, x = 1.355, y = 1.633, width = 0.6345, height = 0.6345) +
  cowplot::draw_plot_label(c("A", "B", "C"), x = c(0, 1.35, 0), y = c(2.25, 2.25, 1.65), size = 25)
mcg_subgraph_with_context
if (!interactive()){
  cowplot::ggsave(plot = mcg_subgraph_with_context,
                  filename = file.path(PROJHOME, "results", "figures", "publication", "mcg-subgraph-with-context.pdf"),
                  width = 8, 
                  height = 10)
}
```
